---
title: "Reproducible research on paper Weibull Regression and Machine Learning Survival Models. Methodology, Comparison and Application to Biomedical Data Related to Cardiac Surgery"
author: "Thalytta Cavalcante, Raydonal Ospina, Víctor Leiva, Xavier Cabezas, Carlos Martin-Barreiro"
date: "`r Sys.Date()`"
output: 
  prettydoc::html_pretty:
    theme: cayman
    highlight: github
vignette: >
  %\VignetteIndexEntry{Vignette Title}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
header-includes:
- \usepackage[english]{babel}
- \usepackage[latin1]{inputenc}
- \usepackage{amsmath,amsthm}
- \usepackage{amssymb,latexsym}
- \usepackage{graphics}
- \usepackage{graphicx}
- \usepackage{amscd,eufrak}
- \usepackage{placeins}
- \usepackage{setspace}
- \usepackage{url}
- \usepackage{color}
- \setcounter{secnumdepth}{-1}
- \usepackage{pgfplots}
- \usepackage{subfigure}
- \usepackage{geometry}
- \usepackage{chngcntr}
- \onehalfspacing
- \counterwithin{figure}{section}
- \counterwithin{table}{section}
---


```{r number_figures, echo=FALSE, warning=FALSE, message=FALSE, eval=TRUE}
suppressPackageStartupMessages(library(knitr))
#https://rstudio-pubs-static.s3.amazonaws.com/274181_55917dc35eee48efac24c0d6fa9f0c11.html
```


```{r startup, echo = FALSE, warning = FALSE, message=FALSE}
knitr::opts_chunk$set(echo = TRUE, 
                      warning = FALSE,
                      message = FALSE,
                      fig.width  = 10,
                      # fig.height = 13,
                      fig.cap = '',
                      fig.align = 'center', 
                      dpi = 72*2)#, global.par = TRUE)

options(width = 150, 
        scipen = 1e3)

# load the package
suppressPackageStartupMessages(library(plyr))
suppressPackageStartupMessages(library(dplyr))
suppressPackageStartupMessages(library(ggplot2))
suppressPackageStartupMessages(library(waffle))
suppressPackageStartupMessages(library(survival))
suppressPackageStartupMessages(library(tidyverse))
suppressPackageStartupMessages(library(MASS))
suppressPackageStartupMessages(library(flexsurv))
suppressPackageStartupMessages(library(dynpred))
suppressPackageStartupMessages(library(readxl))
suppressPackageStartupMessages(library(xtable))
suppressPackageStartupMessages(library(randomForestSRC))
suppressPackageStartupMessages(library(survival))
suppressPackageStartupMessages(library(survminer))
suppressPackageStartupMessages(library(gtsummary))
suppressPackageStartupMessages(library(survival))
suppressPackageStartupMessages(library(lubridate))
suppressPackageStartupMessages(library(ggsurvfit))
suppressPackageStartupMessages(library(gtsummary))
suppressPackageStartupMessages(library(tidycmprsk))
suppressPackageStartupMessages(library(kableExtra))
suppressPackageStartupMessages(library(scales))
suppressPackageStartupMessages(library(ggridges))
suppressPackageStartupMessages(library(reshape))
suppressPackageStartupMessages(library(reshape2))
suppressPackageStartupMessages(library(magrittr))
suppressPackageStartupMessages(library(htmlTable))
suppressPackageStartupMessages(library(gridExtra)) 
suppressPackageStartupMessages(library(Hmisc))
suppressPackageStartupMessages(library(RColorBrewer))
suppressPackageStartupMessages(library(caret))
suppressPackageStartupMessages(library(ggRandomForests))
#=============================================================================
```
# Data set (small view)
```{r, echo=F, message=FALSE, warning=FALSE}
dados <- read_excel("viarapida.xlsm",sheet=1, col_names=T)
colnames(dados)<-c("TCC", "DELTA", "AGE", "PROTOCOL", "RACE", "SEX", "PATT")

dados$PROTOCOL <- as.factor(dados$PROTOCOL)
dados$PROTOCOL <- revalue(dados$PROTOCOL, c("0" = "conventional", 
                                            "1" = "fast-track"))

dados$RACE <- as.factor(dados$RACE)
dados$RACE <- revalue(dados$RACE, c("1" = "white", 
                                    "2" = "black", 
                                    "3" = "asian"))

dados$SEX <- as.factor(dados$SEX)
dados$SEX <- revalue(dados$SEX, c("0" = "female", 
                                  "1" = "male"))

dados$PATT <- as.factor(dados$PATT)
dados$PATT <- revalue(dados$PATT, c("0" = "congenital",
                                                    "1" = "coronary"))
kable(head(dados, n=5))


```

## Table 1
```{r, echo=F, warning=FALSE, message=FALSE, comment=NA}
dt <- dados %>% group_by(PATT, PROTOCOL) %>% summarise(N=n(), Mean=mean(AGE), SD =sd(AGE), Min=min(AGE), Median=median(AGE), Max=max(AGE)) %>% dplyr::mutate_if(is.numeric, round, 1)

tab.AGE <- xtabs(cbind(N,Mean, SD,Min,Median,Max) ~ PATT+PROTOCOL,data = dt) 

ftable(tab.AGE)
```

# Table 2
```{r, echo=F, warning=FALSE, message=FALSE, comment=NA}

dt <- dados %>% group_by(PATT, PROTOCOL, SEX) %>% summarise(N=n()) %>% mutate(Prop=N/sum(N)*100) %>% dplyr::mutate_if(is.numeric, round, 1)

tab.PATT <- xtabs(cbind(N,Prop) ~  PATT+SEX +PROTOCOL, data = dt)

ftable(tab.PATT)
```

# Table 3
```{r, echo=F, warning=FALSE, message=FALSE, comment=NA}

dt <- dados %>% group_by(PATT, PROTOCOL, RACE) %>% summarise(N=n()) %>% mutate(Prop=N/sum(N)*100) %>% dplyr::mutate_if(is.numeric, round, 1)

tab.RACE <- xtabs(cbind(N,Prop) ~  PATT+RACE +PROTOCOL, data = dt)

ftable(tab.RACE)
```

# Figure 3
```{r, echo=F, warning=FALSE, message=FALSE, comment=NA, results='hide'}
# Fig 3a
cols <-  hcl.colors(length(levels(dados$SEX)), "blue-Yellow 3")
tb <-table(dados$SEX)
df <- data.frame(tb)
colnames(df) <- c("sex","count")
# Hole size
hsize <- 1.5
p.1 <- ggplot(df, aes(x = hsize, y = count, fill = sex)) +
  geom_col(color = "black") +
  geom_text(aes(label = paste("", round(count/145*100,0), "%", sep = "")),
            position = position_stack(vjust = 0.5), size=9) +
  coord_polar(theta = "y") +
  scale_fill_manual(values = cols) +
  xlim(c(0.2, hsize + 0.5)) + guides(fill=guide_legend(title="")) + 
  theme(panel.background = element_rect(fill = "white"),
        panel.grid = element_blank(),
        axis.title = element_blank(),
        axis.ticks = element_blank(),
        axis.text = element_blank(), legend.text=element_text(size=rel(1.5)), legend.position="bottom") +
  annotate(geom = 'text', x = 0.25, y = 0.0, label = "Sex", color = 'black', size = 9)

# save the plot
ggsave("Fig3a.pdf", p.1, width = 8, height = 8)

# Fig 3b
cols <-  hcl.colors(length(levels(dados$PROTOCOL)), "Temps")
tb <-table(dados$PROTOCOL)
df <- data.frame(tb)
colnames(df) <- c("protocol","count")
# Hole size
hsize <- 1.5
p.2 <- ggplot(df, aes(x = hsize, y = count, fill = protocol)) +
  geom_col(color = "black") +
  geom_text(aes(label = paste("", round(count/145*100,0), "%", sep = "")),
            position = position_stack(vjust = 0.5), size=9) +
  coord_polar(theta = "y") +
  scale_fill_manual(values = cols) +
  xlim(c(0.2, hsize + 0.5)) + guides(fill=guide_legend(title="")) + 
  theme(panel.background = element_rect(fill = "white"),
        panel.grid = element_blank(),
        axis.title = element_blank(),
        axis.ticks = element_blank(),
        axis.text = element_blank(), legend.text=element_text(size=rel(1.5)), legend.position="bottom") +
  annotate(geom = 'text', x = 0.25, y = 0.0, label = "Protocol Care", color = 'black', size = 9)
# save the plot
ggsave("Fig3b.pdf", p.2, width = 8, height = 8)

# Fig 3c
cols <-  hcl.colors(length(levels(dados$PATT)), "Green-Orange")
tb <-table(dados$PATT)
df <- data.frame(tb)
colnames(df) <- c("type_patient","count")
# Hole size
hsize <- 1.5
p.3 <- ggplot(df, aes(x = hsize, y = count, fill = type_patient)) +
  geom_col(color = "black") +
  geom_text(aes(label = paste("", round(count/145*100,0), "%", sep = "")),
            position = position_stack(vjust = 0.5), size=9) +
  coord_polar(theta = "y") +
  scale_fill_manual(values = cols) +
  xlim(c(0.2, hsize + 0.5)) + guides(fill=guide_legend(title="")) + 
  theme(panel.background = element_rect(fill = "white"),
        panel.grid = element_blank(),
        axis.title = element_blank(),
        axis.ticks = element_blank(),
        axis.text = element_blank(), legend.text=element_text(size=rel(1.5)),legend.position="bottom") +
  annotate(geom = 'text', x = 0.25, y = 0.0, label = "Type patient", color = 'black', size = 9)

ggsave("Fig3c.pdf", p.3, width = 8, height = 8)

out <- grid.arrange(p.1, p.2, p.3,  ncol=3,   widths = c(1.5, 1.5, 1.5))
return(out)
```


# Figure 4
```{r, echo=F, warning=FALSE, message=FALSE, comment=NA, results='hide'}
# Fig 4a
cols <-  hcl.colors(length(levels(dados$PROTOCOL)), "Temps")
p.1 <- ggplot(data = dados, aes(x=PROTOCOL, y=AGE)) +
  geom_dotplot(aes(x = "1"), binaxis="y", binwidth=0.8, stackdir="center") +
  geom_boxplot(aes(x = "2"), width=0.6, fill = cols) +
  facet_wrap(~PROTOCOL, strip.position = "bottom") +
  labs(title="Protocol Care", y = "Age") +
  theme(
    plot.title = element_text(hjust = 0.5, size=22),
    # panel.spacing.x = unit(0, "npc"),
    panel.background = element_rect(fill = "white", colour = "black"),
    strip.background = element_blank(),
    strip.text = element_text(size=20),
    axis.text.x = element_blank(),
    axis.ticks.x = element_blank(),
    axis.title.x = element_blank(),
    text = element_text(size = 20)
  )
ggsave("Fig4a.pdf", p.1, width = 8, height = 8)

# Fig 4b
cols <-  hcl.colors(length(levels(dados$PATT)), "Green-Orange")
p.2 <- ggplot(data = dados, aes(x=PATT, y=AGE)) +
  geom_dotplot(aes(x = "1"), binaxis="y", binwidth=0.8, stackdir="center") +
  geom_boxplot(aes(x = "2"), width=0.6, fill = cols) +
  facet_wrap(~PATT, strip.position = "bottom") +
  labs(title="Type patient", y = "Age") +
  theme(
    plot.title = element_text(hjust = 0.5, size=22),
    # panel.spacing.x = unit(0, "npc"),
    panel.background = element_rect(fill = "white", colour = "black"),
    strip.background = element_blank(),
    strip.text = element_text(size=20),
    axis.text.x = element_blank(),
    axis.ticks.x = element_blank(),
    axis.title.x = element_blank(),
    text = element_text(size = 20)
  )

ggsave("Fig4b.pdf", p.2, width = 8, height = 8)

out <- grid.arrange(p.1, p.2,  ncol=2,   widths = c(1.5, 1.5))
return(out)

```

# Figure 5
### Figure 5a
```{r, echo=F, warning=FALSE, message=FALSE, comment=NA}
cols <-  hcl.colors(length(unique(dados$SEX)), "blue-Yellow 3")

h.temp <- rcorrcens(formula=Surv(TCC,DELTA)~SEX, data=dados)
harrel <- round(h.temp[,"C"], digits =3)


fit <- survfit(Surv(TCC,DELTA)~SEX, data=dados) 

pvalor <- round(surv_pvalue(fit)$pval, digits = 3)


p.1 <- ggsurvplot(
  fit,
  data = dados,
  size = 0.7,                 # change line size
  palette =
   cols,# custom color palettes
  conf.int = TRUE,          # Add confidence interval
  # pval = TRUE,              # Add p-value
  risk.table = TRUE,        # Add risk table
  xlab = "time (in hours) by sex",
  ylab = "S(t)",
  legend.title = "Sex",
 # risk.table.col = "sex",# Risk table color by groups
  # legend.labs =
  #   c("Male", "Female"),    # Change legend labels
  risk.table.height = 0.25, # Useful to change when you have multiple groups
  ggtheme = theme_bw(),      # Change ggplot2 theme
 risk.table.y.text = FALSE,
 tables.theme = theme_cleantable(),
 legend.labs =
   c("Female","Male")    # change legend labels.
)

p.1$plot <- p.1$plot + 
  theme( legend.position = c(0.8, 0.85), #"none", #  "right", #
         legend.direction="vertical",
         legend.title = element_blank()) 

p.1$plot <- p.1$plot+
  annotate("text", x=1.5, y=0.1,
            label = deparse(bquote(~"C-ind" ==~  .(harrel))),
            parse = T)

p.1$plot <- p.1$plot+
  annotate("text", x=1.5, y=0.2,
            label = deparse(bquote(~"p-val" ==~  .(pvalor))),
            parse = T)

p.1

kbl(rbind(c("C-Index","p-val:lograk"), c(harrel,pvalor)))

# add method to grid.draw
grid.draw.ggsurvplot <- function(x){
  survminer:::print.ggsurvplot(x, newpage = FALSE)
}

# Remember to pass object `p`.
ggsave("Fig5a.pdf", plot = p.1, width = 6, height = 6 )
```

### Figure 5b
```{r, echo=F, warning=FALSE, message=FALSE, comment=NA}
cols <-  hcl.colors(length(unique(dados$RACE)))

h.temp <- rcorrcens(formula=Surv(TCC,DELTA)~RACE, data=dados)
harrel <- round(h.temp[,"C"], digits =3)


fit <- survfit(Surv(TCC,DELTA)~RACE, data=dados) 

pvalor <- round(surv_pvalue(fit)$pval, digits = 3)


p.1 <- ggsurvplot(
  fit,
  data = dados,
  size = 0.7,                 # change line size
  palette =
   cols,# custom color palettes
  conf.int = TRUE,          # Add confidence interval
  # pval = TRUE,              # Add p-value
  risk.table = TRUE,        # Add risk table
  xlab = "time (in hours) of stay in the surgery center according to race",
  ylab = "S(t)",
  legend.title = "Race",
 # risk.table.col = "sex",# Risk table color by groups
  # legend.labs =
  #   c("Male", "Female"),    # Change legend labels
  risk.table.height = 0.25, # Useful to change when you have multiple groups
  ggtheme = theme_bw(),      # Change ggplot2 theme
 risk.table.y.text = FALSE,
 tables.theme = theme_cleantable(),
 legend.labs =
  c("white","black","asian")  # change legend labels.
)

p.1$plot <- p.1$plot + 
  theme( legend.position = c(0.8, 0.85), #"none", #  "right", #
         legend.direction="vertical",
         legend.title = element_blank()) 

p.1$plot <- p.1$plot+
  annotate("text", x=1.5, y=0.1,
            label = deparse(bquote(~"C-ind" ==~  .(harrel))),
            parse = T)

p.1$plot <- p.1$plot+
  annotate("text", x=1.5, y=0.2,
            label = deparse(bquote(~"p-val" ==~  .(pvalor))),
            parse = T)

p.1

kbl(rbind(c("C-Index","p-val:lograk"), c(harrel,pvalor)))

# add method to grid.draw
grid.draw.ggsurvplot <- function(x){
  survminer:::print.ggsurvplot(x, newpage = FALSE)
}

# Remember to pass object `p`.
ggsave("Fig5b.pdf", plot = p.1, width = 6, height = 6)
```

### Figure 5c
```{r, echo=F, warning=FALSE, message=FALSE, comment=NA}
cols <-  hcl.colors(length(unique(dados$PATT)))

h.temp <- rcorrcens(formula=Surv(TCC,DELTA)~PATT, data=dados)
harrel <- round(h.temp[,"C"], digits =3)


fit <- survfit(Surv(TCC,DELTA)~PATT, data=dados) 

pvalor <- round(surv_pvalue(fit)$pval, digits = 3)


p.1 <- ggsurvplot(
  fit,
  data = dados,
  size = 0.7,                 # change line size
  palette =
   cols,# custom color palettes
  conf.int = TRUE,          # Add confidence interval
  # pval = TRUE,              # Add p-value
  risk.table = TRUE,        # Add risk table
 xlab="time (in hours) of stay in the surgery center according to the type of patient",
  ylab = "S(t)",
  legend.title = "type",
 # risk.table.col = "sex",# Risk table color by groups
  # legend.labs =
  #   c("Male", "Female"),    # Change legend labels
  risk.table.height = 0.25, # Useful to change when you have multiple groups
  ggtheme = theme_bw(),      # Change ggplot2 theme
 risk.table.y.text = FALSE,
 tables.theme = theme_cleantable(),
 legend.labs =
  c("congenital","coronary")
)

p.1$plot <- p.1$plot + 
  theme( legend.position = c(0.8, 0.85), #"none", #  "right", #
         legend.direction="vertical",
         legend.title = element_blank()) 

p.1$plot <- p.1$plot+
  annotate("text", x=1.9, y=0.1,
            label = deparse(bquote(~"C-ind" ==~  .(harrel))),
            parse = T)

p.1$plot <- p.1$plot+
  annotate("text", x=1.5, y=0.2,
            label = deparse(bquote(~"p-val" ==~  .(pvalor))),
            parse = T)

p.1

kbl(rbind(c("C-Index","p-val:lograk"), c(harrel,pvalor)))

# add method to grid.draw
grid.draw.ggsurvplot <- function(x){
  survminer:::print.ggsurvplot(x, newpage = FALSE)
}

# Remember to pass object `p`.
ggsave("Fig5c.pdf", plot = p.1, width = 6, height = 6)
```

### Figure 5d
```{r, echo=F, warning=FALSE, message=FALSE, comment=NA}
cols <-  hcl.colors(length(unique(dados$PROTOCOL))+1)[-3]

h.temp <- rcorrcens(formula=Surv(TCC,DELTA)~PROTOCOL, data=dados)
harrel <- round(h.temp[,"C"], digits =3)


fit <- survfit(Surv(TCC,DELTA)~PROTOCOL, data=dados) 

pvalor <- round(surv_pvalue(fit)$pval, digits = 3)


p.1 <- ggsurvplot(
  fit,
  data = dados,
  size = 0.7,                 # change line size
  palette =
   cols,# custom color palettes
  conf.int = TRUE,          # Add confidence interval
  # pval = TRUE,              # Add p-value
  risk.table = TRUE,        # Add risk table
  xlab="time (in hours) of stay in the surgery center according to the protocol",
  ylab = "S(t)",
  legend.title = "Protocol",
 # risk.table.col = "sex",# Risk table color by groups
  # legend.labs =
  #   c("Male", "Female"),    # Change legend labels
  risk.table.height = 0.25, # Useful to change when you have multiple groups
  ggtheme = theme_bw(),      # Change ggplot2 theme
 risk.table.y.text = FALSE,
 tables.theme = theme_cleantable(),
 legend.labs =
  c("conventional","fast-track")
)

p.1$plot <- p.1$plot + 
  theme( legend.position = c(0.8, 0.85), #"none", #  "right", #
         legend.direction="vertical",
         legend.title = element_blank()) 

p.1$plot <- p.1$plot+
  annotate("text", x=1.5, y=0.1,
            label = deparse(bquote(~"C-ind" ==~  .(harrel))),
            parse = T)

p.1$plot <- p.1$plot+
  annotate("text", x=1.8, y=0.2,
            label = deparse(bquote(~"p-val" ==~  .(pvalor))),
            parse = T)

p.1

kbl(rbind(c("C-Index","p-val:lograk"), c(harrel,pvalor)))

# add method to grid.draw
grid.draw.ggsurvplot <- function(x){
  survminer:::print.ggsurvplot(x, newpage = FALSE)
}

# Remember to pass object `p`.
ggsave("Fig5d.pdf", plot = p.1, width = 6, height = 6)
```

# Weibull regression model parameter estimates
### Table 4 - All covariates
```{r, echo=F, warning=FALSE, message=FALSE, comment=NA}

ajuste.W<-survreg(Surv(TCC,DELTA)~AGE+PROTOCOL+RACE+SEX+PATT, dist="weibull", data=dados)
summary(ajuste.W)
```

### Table 5 - After model selection
```{r, echo=F, warning=FALSE, message=FALSE, comment=NA}
ajuste.w1<-survreg(Surv(TCC,DELTA)~PROTOCOL+RACE+PATT, dist="weibull", data=dados)
summary(ajuste.w1)
```

# Goodness-of-fit
## Figure 6
### Figure 6a - goodness-of-fit
```{r, echo=F, warning=FALSE, message=FALSE, comment=NA}
xb<-ajuste.w1$linear.predictors
gama.hat<-1/ajuste.w1$scale

res.cox.snell <-numeric(145)
for (i in 1:length(dados$TCC)) {
	if(dados$DELTA[i]==1){
	res.cox.snell[i]=(dados$TCC[i]*exp(-xb[i]))^gama.hat
      } else {
      	res.cox.snell[i]=((dados$TCC[i]*exp(-xb[i]))^gama.hat) + 1
	}
}

ekm<-survfit(Surv(res.cox.snell,DELTA)~1, type=c("kaplan-meier"), data=dados)

dados$xb <- xb 
dados$res.cox.snell <- res.cox.snell
db.1 <- data.frame(time = ekm$time, surv=ekm$surv, sexp1=exp(-ekm$time))

p <- ggplot()+ geom_point(aes(x=surv, y=sexp1), db.1, col="black")+theme_bw()
p <- p+labs(title="", y = "S(res) : Standard exponential", x="S(res): Kaplan-Meier")+geom_abline(intercept = 0, slope = 1, col="red")+xlim(0,1)+ylim(0,1)

p

ggsave("Fig6a.pdf", plot = p, width = 5, height = 5)
```

### Figure 6b - residual
```{r, echo=F, warning=FALSE, message=FALSE, comment=NA}

res <- sort(dados$res.cox.snell) 
exp1 <- exp(-res)
db.2 <- data.frame(res=res, exp1=exp1)
p <- ggplot()+ geom_line(aes(x=time, y=surv, color="l1", linetype="l1"), db.1)+geom_line(aes(x=res, y=exp1, color="l2", linetype="l2"), db.2)+
  scale_colour_manual(name = " ", values = c("l1"="black", "l2"="red"), 
                      labels = c("Kaplan-Meier", "Standard exponential"))+
  scale_linetype_manual(name = " ", values = c("l1"="solid", "l2"="solid"), 
                        labels = c("Kaplan-Meier", "Standard exponential"))

p <- p+labs(title="", y = "S(res) estimated", x="Residuals")+theme_bw()+theme(
    legend.position = c(0.75, 0.85)
    # ,
    # legend.justification = c("left", "top"),
    # legend.box.just = "left"
  )

p 
ggsave("Fig6b.pdf", plot = p, width = 5, height = 5)
```

### Figure 6c
```{r, echo=F, warning=FALSE, message=FALSE, comment=NA}
dados$martingale <- dados$DELTA - dados$res.cox.snell
dados$index <-1:nrow(dados) 
dados$Name <- as.character(dados$index)


p <- ggplot(data=dados)+ geom_point(aes(x=index, y=martingale), col="black")+ geom_text(aes(x=index, y=martingale, label=ifelse(martingale > -4,'', as.character(index))),hjust=-0.5,vjust=0)+theme_bw()
p <- p+labs(title="", y = "Residuals", x="Index")

p
ggsave("Fig6c.pdf", plot = p, width = 5, height = 5)
```

### Figure 6d
```{r, echo=F, warning=FALSE, message=FALSE, comment=NA}
dados$res.martingale <- dados$DELTA - dados$res.cox.snell
dados$index <-1:nrow(dados) 
dados$res.deviance <- sign(dados$res.martingale)*(-2*(dados$res.martingale+dados$DELTA*log(dados$DELTA-dados$res.martingale)))^(1/2)

p <- ggplot(data=dados)+ geom_point(aes(x=index, y=res.deviance), col="black")+ geom_text(aes(x=index, y=res.deviance, label=ifelse(res.deviance > -2.5,'', as.character(index))),hjust=-0.5,vjust=0)+theme_bw()
p <- p+labs(title="", y = "Residuals", x="Index")

p
ggsave("Fig6d.pdf", plot = p, width = 5, height = 5)
```

### Figure 6e

```{r, echo=F, warning=FALSE, message=FALSE, comment=NA}


dados$corxb <- as.factor(xb)

p <- ggplot(data=dados)+ geom_point(aes(x=xb, y=res.martingale, color=corxb))+ scale_color_brewer(palette="Dark2")+geom_text(aes(x=xb, y=res.martingale, label=ifelse( (res.martingale< -4)  ,as.character(index),'' )),hjust=1.3,vjust=0.3)+theme_bw()
p <- p+labs(title="", y = "Residuals", x="Adjusted values")
p <- p+  annotate(geom = 'text', x =dados$xb[124], y =dados$res.martingale[124], label = "124", color = 'black', hjust=1.3,vjust=0.3)
p <- p+  annotate(geom = 'text', x =dados$xb[144], y =dados$res.martingale[144], label = "144", color = 'black', hjust=1.3,vjust=0.3)+theme(legend.position="none")

p
ggsave("Fig6e.pdf", plot = p, width = 5, height = 5)
```

### Figure 6f
```{r, echo=F, warning=FALSE, message=FALSE, comment=NA}


dados$corxb <- as.factor(xb)

p <- ggplot(data=dados)+ geom_point(aes(x=xb, y=res.deviance, color=corxb))+ scale_color_brewer(palette="Dark2")+geom_text(aes(x=xb, y=res.deviance, label=ifelse( (res.deviance < -2.6)  ,as.character(index),'' )),hjust=1.3,vjust=0.3)+theme_bw()
p <- p+labs(title="", y = "Residuals", x="Adjusted values")
p <- p+  annotate(geom = 'text', x =dados$xb[124], y =dados$res.deviance[124], label = "124", color = 'black', hjust=1.3,vjust=0.3)
p <- p+  annotate(geom = 'text', x =dados$xb[144], y =dados$res.deviance[144], label = "144", color = 'black', hjust=1.3,vjust=0.3)+theme(legend.position="none")

p
ggsave("Fig6f.pdf", plot = p, width = 5, height = 5)
```

# Influence analysis of points in Weibull model
## Models after remove the influential observation

### Remove observation 64

```{r, echo=F, warning=FALSE, message=FALSE, comment=NA}
modelo.64<-survreg(Surv(TCC,DELTA)~PROTOCOL+PATT, subset=-64, dist="weibull", data=dados)
summary(modelo.64)
```

### Remove observation 101

```{r, echo=F, warning=FALSE, message=FALSE, comment=NA}
modelo.101<-survreg(Surv(TCC,DELTA)~PROTOCOL+PATT, subset=-101, dist="weibull", data=dados)
summary(modelo.101)
```

### Remove observation 124

```{r, echo=F, warning=FALSE, message=FALSE, comment=NA}
modelo.124<-survreg(Surv(TCC,DELTA)~PROTOCOL+PATT, subset=-124, dist="weibull", data=dados)
summary(modelo.124)
```


### Remove observation 144

```{r, echo=F, warning=FALSE, message=FALSE, comment=NA}
modelo.144<-survreg(Surv(TCC,DELTA)~PROTOCOL+PATT, subset=-144, dist="weibull", data=dados)
summary(modelo.144)
```

### Remove the set of observations (63, 101,  144) 

```{r, echo=F, warning=FALSE, message=FALSE, comment=NA}
modelo.64.101.144<-survreg(Surv(TCC,DELTA)~PROTOCOL+PATT, subset=c(-64,-101,-124,-144), dist="weibull", data=dados)
summary(modelo.64.101.144)
```


### Remove the set of observations (63, 101, 124, 144) 

```{r, echo=F, warning=FALSE, message=FALSE, comment=NA}
modelo.64.101.124.144<-survreg(Surv(TCC,DELTA)~PROTOCOL+PATT, subset=c(-64,-101,-124,-144), dist="weibull", data=dados)
summary(modelo.64.101.124.144)
```

### Table 7 - After model selection
```{r, echo=F, warning=FALSE, message=FALSE, comment=NA}
ajuste.w1<-survreg(Surv(TCC,DELTA)~PROTOCOL+PATT, dist="weibull", data=dados)
summary(ajuste.w1)

temp <- concordance(ajuste.w1) # Calcula Índice C

a <- temp$concordance
b <- 1-a
kable(rbind(c("C-index","error-rate"), c(a,b)), caption="Performance:  Weibull model", digits = 2) 
```

# Goodness-of-fit
## Figure 7
### Figure 7a - goodness-of-fit
```{r, echo=F, warning=FALSE, message=FALSE, comment=NA}
xb<-ajuste.w1$linear.predictors
gama.hat<-1/ajuste.w1$scale

res.cox.snell <-numeric(145)
for (i in 1:length(dados$TCC)) {
	if(dados$DELTA[i]==1){
	res.cox.snell[i]=(dados$TCC[i]*exp(-xb[i]))^gama.hat
      } else {
      	res.cox.snell[i]=((dados$TCC[i]*exp(-xb[i]))^gama.hat) + 1
	}
}

ekm<-survfit(Surv(res.cox.snell,DELTA)~1, type=c("kaplan-meier"), data=dados)

dados$xb <- xb 
dados$res.cox.snell <- res.cox.snell
db.1 <- data.frame(time = ekm$time, surv=ekm$surv, sexp1=exp(-ekm$time))

p <- ggplot()+ geom_point(aes(x=surv, y=sexp1), db.1, col="black")+theme_bw()
p <- p+labs(title="", y = "S(res) : Standard exponential", x="S(res): Kaplan-Meier")+geom_abline(intercept = 0, slope = 1, col="red")+xlim(0,1)+ylim(0,1)

p

ggsave("Fig7a.pdf", plot = p, width = 5, height = 5)
```

### Figure 7b - residual
```{r, echo=F, warning=FALSE, message=FALSE, comment=NA}

res <- sort(dados$res.cox.snell) 
exp1 <- exp(-res)
db.2 <- data.frame(res=res, exp1=exp1)
p <- ggplot()+ geom_line(aes(x=time, y=surv, color="l1", linetype="l1"), db.1)+geom_line(aes(x=res, y=exp1, color="l2", linetype="l2"), db.2)+
  scale_colour_manual(name = " ", values = c("l1"="black", "l2"="red"), 
                      labels = c("Kaplan-Meier", "Standard exponential"))+
  scale_linetype_manual(name = " ", values = c("l1"="solid", "l2"="solid"), 
                        labels = c("Kaplan-Meier", "Standard exponential"))

p <- p+labs(title="", y = "S(res) estimated", x="Residuals")+theme_bw()+theme(
    legend.position = c(0.75, 0.85)
    # ,
    # legend.justification = c("left", "top"),
    # legend.box.just = "left"
  )

p 
ggsave("Fig7b.pdf", plot = p, width = 5, height = 5)
```

### Figure 7c
```{r, echo=F, warning=FALSE, message=FALSE, comment=NA}
dados$martingale <- dados$DELTA - dados$res.cox.snell
dados$index <-1:nrow(dados) 
dados$Name <- as.character(dados$index)


p <- ggplot(data=dados)+ geom_point(aes(x=index, y=martingale), col="black")+ geom_text(aes(x=index, y=martingale, label=ifelse(martingale > -4,'', as.character(index))),hjust=-0.5,vjust=0)+theme_bw()
p <- p+labs(title="", y = "Residuals", x="Index")

p
ggsave("Fig7c.pdf", plot = p, width = 5, height = 5)
```

### Figure 7d
```{r, echo=F, warning=FALSE, message=FALSE, comment=NA}
dados$res.martingale <- dados$DELTA - dados$res.cox.snell
dados$index <-1:nrow(dados) 
dados$res.deviance <- sign(dados$res.martingale)*(-2*(dados$res.martingale+dados$DELTA*log(dados$DELTA-dados$res.martingale)))^(1/2)

p <- ggplot(data=dados)+ geom_point(aes(x=index, y=res.deviance), col="black")+ geom_text(aes(x=index, y=res.deviance, label=ifelse(res.deviance > -2.5,'', as.character(index))),hjust=-0.5,vjust=0)+theme_bw()
p <- p+labs(title="", y = "Residuals", x="Index")

p
ggsave("Fig7d.pdf", plot = p, width = 5, height = 5)
```

### Figure 7e

```{r, echo=F, warning=FALSE, message=FALSE, comment=NA}


dados$corxb <- as.factor(xb)

p <- ggplot(data=dados)+ geom_point(aes(x=xb, y=res.martingale, color=corxb))+ scale_color_brewer(palette="Dark2")+geom_text(aes(x=xb, y=res.martingale, label=ifelse( (res.martingale< -4)  ,as.character(index),'' )),hjust=1.3,vjust=0.3)+theme_bw()
p <- p+labs(title="", y = "Residuals", x="Adjusted values")
# p <- p+  annotate(geom = 'text', x =dados$xb[124], y =dados$res.martingale[124], label = "124", color = 'black', hjust=1.3,vjust=0.3)
p <- p+  annotate(geom = 'text', x =dados$xb[144], y =dados$res.martingale[144], label = "144", color = 'black', hjust=1.3,vjust=0.3)+theme(legend.position="none")

p
ggsave("Fig7e.pdf", plot = p, width = 5, height = 5)
```

### Figure 7f
```{r, echo=F, warning=FALSE, message=FALSE, comment=NA}


dados$corxb <- as.factor(xb)

p <- ggplot(data=dados)+ geom_point(aes(x=xb, y=res.deviance, color=corxb))+ scale_color_brewer(palette="Dark2")+geom_text(aes(x=xb, y=res.deviance, label=ifelse( (res.deviance < -2.5)  ,as.character(index),'' )),hjust=1.3,vjust=0.3)+theme_bw()
p <- p+labs(title="", y = "Residuals", x="Adjusted values")
# p <- p+  annotate(geom = 'text', x =dados$xb[124], y =dados$res.deviance[124], label = "124", color = 'black', hjust=1.3,vjust=0.3)
p <- p+  annotate(geom = 'text', x =dados$xb[144], y =dados$res.deviance[144], label = "144", color = 'black', hjust=1.3,vjust=0.3)+theme(legend.position="none")

p
ggsave("Fig7f.pdf", plot = p, width = 5, height = 5)
```


# Random Forest result

```{r, echo=F, warning=FALSE, message=FALSE, comment=NA}
rm(dados)
# clean memory and data set to RF
gc()

dados <- read_excel("viarapida.xlsm",sheet=1, col_names=T)
colnames(dados)<-c("TCC", "DELTA", "AGE", "PROTOCOL", "RACE", "SEX", "PATT")
dados <- na.omit(dados)

NTREE = 1000 # N?mero de arboles.
NODESIZE = 10 # Tamaño minimo del nodo terminal. 
## Los valores predeterminados son: supervivencia (15), 
## riesgo competitivo (15), regresión (5), clasificación (1), 
## resultados mixtos (3), sin supervisión (3). 
## Se recomienda experimentar con diferentes valores.
MTRY = 3 # Número de variables a dividir posiblemente en cada nodo.
NSPLIT = 0 # Entero no negativo que especifica el n?mero de divisiones 
# aleatorias para dividir una variable. Cuando es cero, 
# se utilizan todos los valores de divisi?n (división determinista), 
# que puede ser más lento. De manera predeterminada, se usa 10, 
# excepto para regresión y clasificación, 
# que usa cero (para división determinista).
IMPORTANCE = TRUE

set.seed(456) ## 1116
#set.seed(3456)
trainIndex <- createDataPartition(dados$DELTA, p = .8, 
                                  list = FALSE, 
                                  times = 1)

train <- dados[ trainIndex,]
test  <- dados[-trainIndex,]


obj <- rfsrc(Surv(TCC,DELTA)~., train,
             ntree = NTREE, 
             nodesize = NODESIZE, 
             mtry = MTRY, 
             nsplit = NSPLIT, 
             importance = IMPORTANCE, seed=435) # 6650  105 6543 (este funciona)
```

```{r, echo=F, warning=FALSE, message=FALSE, comment=NA}
p <- plot(gg_error(obj)) + geom_line(size=0.75) +theme_bw()

p 
ggsave("Fig8a.pdf", plot = p, width = 5, height = 5)


p <- plot(gg_vimp(obj)) + scale_size_area() + ylab("Variable importance") +theme_bw()

p
ggsave("Fig8b.pdf", plot = p, width = 5, height = 5)

```

# Table 9

```{r, echo=F, warning=FALSE, message=FALSE, comment=NA}
a <- 1-get.cindex(time = obj$yvar[,1], 
             censoring = obj$yvar[,2], 
             predicted = obj$predicted.oob)

b <- 1-a
kable(rbind(c("C-index","error-rate"), c(a,b)), caption="Performance:  RF(training)", digits = 2) 

cat("Variable Importance")
vimp(obj, importance = "random")
vimp(obj, importance = "random")$importance

pred <- predict(obj, test,  na.action = "na.impute")
a <- 1-get.cindex(time = pred$yvar[,1], 
             censoring = pred$yvar[,2], 
             predicted = pred$predicted)
b <- 1-a
kable(rbind(c("C-index","error-rate"), c(a,b)), caption="Performance:  RF(test)", digits = 2) 


pred

#plot(pred)
#plot.survival(pred)


```


